---
title: "Visualizing the Music of Tom Petty and the Heartbreakers Using R"
author: Sadie Kratt
subtitle: GEO 511
---

# Introduction

Tom Petty is one of the greatest rock and roll artists of the 20th century. He tragically died in 2017 right after he finished up his final tour. Because he is my favorite artist, I have decided to honor him by creating a data visualization of his work through R packages.

Problems to be addressed:
* How many studio albums did he release and when?
* What is the danceability of his songs?
* How positive are his songs?
* What are the top 10 hits of his career?
* Where has the band toured?
* What are the top 5 places they have toured?

# Materials and methods

[~ 200 words]

Narrative: Clear narrative description of the data sources and methods. Includes data from at least two sources that were integrated / merged in R.

Code: The code associated with the project is well organized and easy to follow. Demonstrates mastery of R graphics and functions.

Data: The underlying data are publicly accessible via the web and downloaded/accessed within the Rmd script. If you want to use your own data, you must make it available on a website (e.g. Figshare) so that others are able to re-run your code.

You can do bullets like this:

* The first most important thing
* The second most important thing
* The third most important thing

You can do numbers like this:

1. The first most important thing
2. The second most important thing
3. The third most important thing

See [http://rmarkdown.rstudio.com/](http://rmarkdown.rstudio.com/) for all the amazing things you can do.


Loading the Spotify API
```{r}
#Load Spotify package and get access
library(spotifyr)
Sys.setenv(SPOTIFY_CLIENT_ID = '097f7729a69e451da9624441ccdb54bd')
Sys.setenv(SPOTIFY_CLIENT_SECRET = 'd1b68552cb2843929b58db9b964f87e9')

access_token <- get_spotify_access_token()
```

Load Required Packages

```{r, message=F, warning=F}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(dplyr)
library(knitr)
library(purrr)
library (ggmap)
library(leaflet)
library(kableExtra)
knitr::opts_chunk$set(cache=TRUE)  # cache the results for quick compiling
```

## Download and clean all required data
```{r}
n=20
data=data.frame(x=runif(n,-180,180),
                y=runif(n,-60,60),
                size = runif(n, 5, 20),
                category = factor(
                  sample(letters[1:5], n, replace = TRUE)
                  ),
                value = rnorm(n))
```

```{r, results='asis'}
data %>% 
  slice(1:10) %>% #show only 1:n rows
  kable(digits=2,align="c")%>% #make table and round to two digits
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed", "responsive")) #apply other formatting
```

Add any additional processing steps here.

# Results

[~200 words]

Tables and figures (maps and other graphics) are carefully planned to convey the results of your analysis. Intense exploration and evidence of many trials and failures. The author looked at the data in many different ways before coming to the final presentation of the data.

Show tables, plots, etc. and describe them.

```{r, fig.width=6, fig.height=3, fig.cap="Map of completely random data"}
m <- leaflet(data) %>% 
  addTiles() %>% 
  addCircleMarkers(~x, ~y, radius = ~size,color = ~as.factor(category)) %>% 
  addPopups(~x[2], ~y[2], "Random popup")
m  # a map with the default OSM tile layer
```


```{r}
data %>% 
  ggplot(aes(x=x,y=y,col=category))+
  geom_point()
```

# Conclusions

[~200 words]

Clear summary adequately describing the results and putting them in context. Discussion of further questions and ways to continue investigation.

# References

https://www.thepettyarchives.com/archives/miscellany/performances/setlists
https://www.rdocumentation.org/packages/spotifyr/versions/2.2.1
https://github.com/thomasp85/ggraph

