---
title: "Visualizing the Music of Tom Petty and the Heartbreakers Using R"
author: Sadie Kratt
subtitle: GEO 511
---

# Introduction

Tom Petty is one of the greatest rock and roll artists of the 20th century. He tragically died in 2017 right after he finished up his final tour. Because he is my favorite artist, I have decided to honor him by creating a data visualization of his work through R packages.

Problems to be addressed:
* How many studio albums did he release and when?
* What is the danceability of his songs?
* How positive are his songs?
* What are the top 10 hits of his career?
* Where has the band toured?
* What are the top 5 places they have toured?

# Materials and methods


Loading the Spotify API

```{r}
#Load Spotify package and get access
library(spotifyr)
Sys.setenv(SPOTIFY_CLIENT_ID = '097f7729a69e451da9624441ccdb54bd')
Sys.setenv(SPOTIFY_CLIENT_SECRET = 'd1b68552cb2843929b58db9b964f87e9')

access_token <- get_spotify_access_token()
```

Load Required Packages

```{r, message=F, warning=F}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(dplyr)
library(knitr)
library(purrr)
library (ggmap)
library(leaflet)
library(kableExtra)
knitr::opts_chunk$set(cache=TRUE)  # cache the results for quick compiling
```

## Download and clean all required data
```{r}
#load touring data
touring_data <-read.csv('data/locations.csv')
view(touring_data)

#load spotify music data
#Find artist id to access artist information
tom_petty <- get_artist_audio_features("Tom Petty and the Heartbreakers")
View(tom_petty)

#Load album dataset
petty_albums <- get_artist_albums(
  id= "4tX2TplrkIP4v05BNC903e",
  include_groups = c("album", "single", "appears_on", "compilation"),
  market = NULL,
  limit = 20,
  offset = 0,
  authorization = get_spotify_access_token(),
  include_meta_info = FALSE
)
```

```{r, results='asis'}

```

Add any additional processing steps here.

# Results

[~200 words]

Tables and figures (maps and other graphics) are carefully planned to convey the results of your analysis. Intense exploration and evidence of many trials and failures. The author looked at the data in many different ways before coming to the final presentation of the data.

Show tables, plots, etc. and describe them.



# Conclusions

[~200 words]

Clear summary adequately describing the results and putting them in context. Discussion of further questions and ways to continue investigation.

# References

https://www.thepettyarchives.com/archives/miscellany/performances/setlists
https://www.rdocumentation.org/packages/spotifyr/versions/2.2.1 
https://github.com/thomasp85/ggraph

